---
title: "Project 2 EDA"
author: "Jasmine Coleman"
date: "April 11, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

####Downloading and importing the data
```{r Data}
# getwd()

#Downloading the data
# install.packages("downloader")
# library(downloader)
# download("https://raw.githubusercontent.com/BivinSadler/MSDS-6306-Doing-Data-Science/master/UNIT%2014/CaseStudy2-data.csv","Raw Data/Case Study 2.csv")
# download("https://raw.githubusercontent.com/BivinSadler/MSDS-6306-Doing-Data-Science/master/UNIT%2014/CaseStudy2Validation%20No%20Attrition.csv", "Raw Data/Validation Data.csv")
# list.files()

#Importing the downloaded data
EmpInfo <-read.csv("Raw Data/Case Study 2.csv", header = TRUE, sep = ",")
head(EmpInfo)
str(EmpInfo)
```

##Initial exploratory plots
```{r Emp1}
# install.packages("tidyverse")
# install.packages("dplyr")
# install.packages("purrr")
library(tidyverse)
library(dplyr)
library(ggplot2)
library(purrr)

#Bar chart of factor variables
EmpInfo %>%
  keep(is.factor) %>%
  gather() %>%
  ggplot(aes(value)) +
  facet_wrap(~ key, scales="free") +
  geom_bar() +
  theme(axis.text.x=element_text(size=7, angle=45, vjust=1, hjust=1),
        axis.text.y=element_text(size=6))

#HIstogram of integer variables
EmpInfo %>%
  keep(is.integer) %>%
  gather() %>%
  ggplot(aes(value)) +
  facet_wrap(~ key, scales="free") +
  geom_histogram() +
  theme(axis.text.x=element_text(size=7, angle=45, vjust=1, hjust=1),
        axis.text.y=element_text(size=6))
```

##Cleaning Data Frame  

####The following variables can be removed due to lack of variance (same answer recorded for each record)
- Over18 (Y)
- EmployeeCount (1)
- StandanrHours (80)

####Several variables classified as integars could also be seen as categorical and will be converted to factor types
```{r CleanEmp1}
##Removing Over18 variable since only 1 level of factor & EmployeeCount & StandardHours
EmpInfo2 <- EmpInfo[,-c(10, 23, 28)]
head(EmpInfo2)
str(EmpInfo2)
#Casting Rating Variables as Factors
EmpInfo2$Education <- as.factor(EmpInfo2$Education)
EmpInfo2$EnvironmentSatisfaction <- as.factor(EmpInfo2$EnvironmentSatisfaction)
EmpInfo2$JobInvolvement <- as.factor(EmpInfo2$JobInvolvement)
EmpInfo2$JobLevel <- as.factor(EmpInfo2$JobLevel)
EmpInfo2$JobSatisfaction <- as.factor(EmpInfo2$JobSatisfaction)
EmpInfo2$PerformanceRating <- as.factor(EmpInfo2$PerformanceRating)
EmpInfo2$RelationshipSatisfaction <- as.factor(EmpInfo2$RelationshipSatisfaction)
EmpInfo2$StockOptionLevel <- as.factor(EmpInfo2$StockOptionLevel)
EmpInfo2$TrainingTimesLastYear <- as.factor(EmpInfo2$TrainingTimesLastYear)
EmpInfo2$WorkLifeBalance <- as.factor(EmpInfo2$WorkLifeBalance)

#Checking for NAs
colSums(is.na(EmpInfo2))
```

## Grouping Categorical vs Other Variables
```{r Emp2Stats}
#Grouping Categorical Variables
EmpFactors <- EmpInfo2[,c("Attrition", "BusinessTravel", "Department", "EducationField", "Gender", "JobRole", "MaritalStatus", "OverTime", "Education", "EnvironmentSatisfaction", "JobInvolvement", "JobLevel", "JobSatisfaction", "PerformanceRating", "RelationshipSatisfaction", "StockOptionLevel", "TrainingTimesLastYear", "WorkLifeBalance")]

# EmpInfo2 %>% keep(is.factor)

#Grouping Integer Variables
EmpNumeric <- EmpInfo2 %>% keep(is.integer)

#summary statistics
mysummary<-function(x){
  result<-c(length(x),mean(x),sd(x),sd(x)/length(x),min(x),max(x),IQR(x))
  names(result)<-c("N","Mean","SD","SE","Min","Max","IQR")
  return(result)
}

EmpNumStats <- round(sapply(EmpNumeric, mysummary), 2)
EmpNumStats
EmpFactStats <- sapply(sapply(EmpFactors, table), prop.table)
EmpFactStats <- sapply(EmpFactStats, signif)
EmpFactStats
```

##Correlation Plots
```{r Correlations}
#Correlation Plot of Integer Variables
# install.packages("corrplot")  
library(corrplot)

EmpCorrelations <- cor(EmpInfo2 %>% keep(is.integer))
EmpCorrelationsAll <- cor(EmpInfo[,-c(10, 23, 28)] %>% keep(is.integer))
corrplot(EmpCorrelations, type = "lower", diag = TRUE)
corrplot(EmpCorrelationsAll, type = "lower", diag = TRUE)
```

##Attrition Frequencies
####Way to do this enmasse? (apply variables?)
#####Tables below look at Attrition by Job Department
```{r Attrition}
#Attrition by variable
DeptAttr <- table(EmpInfo2$Department,EmpInfo2$Attrition)
margin.table(DeptAttr, 1) # Department Totals 
margin.table(DeptAttr, 2) # Attrition Totals

prop.table(DeptAttr) # cell percentages
prop.table(DeptAttr, 1) # row percentages 
prop.table(DeptAttr, 2) # column percentages
```

##Additional Analysis
```{r Extra}
# # list of variable names
# variables <- sort(names(EmpInfo2))
# 
# 
# #KNN
# # install.packages("caret")
# library(caret)
# 
# #creating training and test set
# train_perc <- .6
# train_indices <- sample(seq(1,nrow(EmpInfo2),by = 1), train_perc*nrow(EmpInfo2))
# train <- EmpInfo2[train_indices,]
# test <- EmpInfo2[-train_indices,]
# 
# #k=3 Confusion Matrix
# class::knn(train[,c()], test[,-3], train$Attrition, k=3)
# confusionMatrix
```